\documentclass[10pt,twocolumn]{article}

\usepackage[letterpaper,top=0.75in,bottom=0.85in,left=0.75in,right=0.75in]{geometry}
\usepackage{times}
\usepackage{microtype}
\usepackage{booktabs}
\usepackage{enumitem}
\usepackage{listings}
\usepackage{xcolor}
\usepackage{hyperref}

\setlength{\columnsep}{0.25in}
\setlength{\parskip}{0.2em}
\setlength{\parindent}{1em}
\setlist{nosep,leftmargin=*}
\lstset{
  basicstyle=\ttfamily\footnotesize,
  columns=fullflexible,
  breaklines=true,
  frame=single,
  xleftmargin=0.2em,
  xrightmargin=0.2em
}

\input{../results/paper_numbers}

\newcommand{\tool}{KernelScript}

\title{\vspace{-1.0em}\tool: Exploring a Typed Unified-Source Programming Model for eBPF}
\author{Anonymous Authors}
\date{}

\begin{document}
\twocolumn[
\begin{center}
\vspace{-1.0em}
{\LARGE\bfseries \tool: Exploring a Typed Unified-Source Programming Model for eBPF\par}
\vspace{0.6em}
{\large Anonymous Authors\par}
\vspace{0.7em}
\begin{minipage}{0.92\textwidth}
\small
\noindent\textbf{Abstract.}
Realistic eBPF applications span verifier-constrained kernel programs,
userspace loaders, shared maps, skeletons, and sometimes kernel-module code.
This paper studies \tool, a beta domain-specific language (DSL) that explores
whether these pieces can be represented in one type-checked source model and
lowered to conventional libbpf-compatible artifacts.  The scientific question
is whether such a unified model can expose cross-boundary eBPF structure to the
compiler without abandoning the ordinary Linux BPF toolchain.

At commit \KSCommitShort{}, \KSKSCompileOK{} of \KSTotalExamples{} repository
examples compile from \tool{} source and \KSMakeOK{} of \KSTotalExamples{}
build into generated C/eBPF projects.  A source-only scan of
\KSExternalCorpusRepos{} pinned external eBPF repositories covers
\KSExternalCorpusFiles{} files and observes markers for Express Data Path (XDP),
traffic control (TC), tracing, maps, ring buffers, struct\_ops, and
scheduler-extension code in the selected source paths, although it does not
translate or build those applications.
A manually ported external XDP portfolio from \texttt{xdp-tutorial} then
compiles from \tool{}: \KSExternalPortWorkloads{} pinned workloads build to
eBPF, attach, and pass traffic in both \tool{} and original external C/eBPF
form over \KSExternalPortTrials{} local trials, and the map-counter workload
also increments the same XDP\_PASS map key in both variants.
The checked-in evaluation then tests selected early rejections, strict verifier
loading, isolated XDP attachment, matched XDP and TC traffic, perf-event
counters, ring-buffer event delivery, and tcp-congestion struct\_ops
compatibility.  \KSVerifierBuildLoadOK{} of
\KSVerifierBuildObjects{} build-success generated objects verifier-load under a
strict pinned-program criterion, and \KSAttachOK{} of \KSAttachEligible{}
verifier-clean single-section XDP objects attach and detach in isolated
namespaces.  For struct\_ops, shared libbpf runners distinguish direct object
load/attach/detach, loopback TCP byte-transfer behavior, callback reachability
under clean and loss-trigger profiles, and a local skeleton-header repair that
builds \KSStructOpsRepairBuildOK{}/\KSStructOpsRepairExamples{} affected
generated userspace projects.  Scheduler-extension experiments add evidence for
verifier loading and bounded attachment: both the hand-written C/eBPF object and
the generated object verifier-load and pin programs, and an opt-in harness
registers both toy schedulers, runs five bounded progress trials, and
unregisters them cleanly.
The results support an artifact/prototype
claim: \tool{} centralizes recurring project structure for this repository
corpus and exposes concrete compatibility limits, but it does not yet establish
broad runtime equivalence or broad struct\_ops portability.
\end{minipage}
\vspace{1.0em}
\end{center}
]

\section{Introduction}

eBPF gives Linux developers a controlled way to execute application-specific
logic in the kernel~\cite{linux-ebpf}.  Linux developers now use it for packet
processing, traffic control, tracing, profiling, security policy, and scheduler
experimentation.  The
programming model remains difficult because the artifact boundary is split:
an eBPF program is written under verifier constraints, a userspace loader must
open and attach it, maps provide shared state, BPF Type Format (BTF) describes
kernel types, and newer interfaces such as kfuncs and struct\_ops require close
coordination between generated skeleton code and kernel headers.

The usual C/libbpf workflow is powerful but verbose~\cite{libbpf}.  The developer writes
kernel-side C, userspace C, Makefiles, pinning logic, map setup, attachment
code, and cleanup paths.  Libraries in Rust, Go, or Python improve parts of the
workflow~\cite{aya,cilium-ebpf}, but they usually preserve the split between the eBPF program and the
controller.  Tracing languages such as bpftrace provide a higher-level
experience, but intentionally narrow the target domain.  The systems question
is therefore not whether eBPF needs another surface syntax.  The question is
whether a language can encode cross-boundary eBPF application structure in a
compiler-checkable way that remains compatible with the existing kernel
toolchain.

\subsection{Why a Unified Source Model?}

A single source file is not automatically better than multiple files.  The
research value comes from making cross-file invariants visible to the compiler.
For example, a map declaration is simultaneously a kernel data structure, a
userspace file descriptor lookup, a pinning decision, and a type contract
between programs.  A program reference is simultaneously a source-level
function name, a generated BPF program section, a skeleton field, a file
descriptor, and possibly a BPF link.  In C/libbpf, these relationships are
implicit naming conventions and generated-header dependencies.  In \tool{},
they can be first-class language values.  The same idea also applies to
cross-language handoff: a userspace stage can \texttt{exec()} into a Python
script while inheriting already-open map file descriptors, and the compiler can
generate the Python-side wrapper from the same source declarations.  In that
case, the shared map schema is not restated in ad hoc FFI glue; it remains part
of the unified source model.

This observation motivates the paper's main scientific question: can a typed
unified-source model make cross-boundary eBPF application structure explicit to
the compiler while preserving the ordinary Linux BPF toolchain path?  A tracing-only language such as
bpftrace~\cite{bpftrace} can be concise by excluding XDP, TC, struct\_ops, and
kfunc workflows.  A host-language binding can be ergonomic by focusing on
userspace loading.  \tool{} attempts a harder point in the design space:
preserve multiple eBPF program types and still give the compiler enough
structure to generate the loader, maps, and build logic.

\tool{} is a recent DSL that makes this question concrete.  It lets developers
write eBPF entry points, helper functions, userspace control flow, maps,
tail-call style delegation, ring buffers, perf events, kfunc declarations, and
kernel-module code in one source language.  The compiler lowers this source to
eBPF C, userspace C, optional module C, a Makefile, and, when a userspace stage
hands off to Python through \texttt{exec()}, an auto-generated Python wrapper
for inherited map descriptors.  The normal libbpf and bpftool path can then
produce objects, skeletons, and binaries.  Because
\tool{} is explicitly beta software, its most useful research role today is as
a prototype for studying the design space rather than as a production-ready
replacement for libbpf.

This paper makes three contributions.  First, it formulates \tool{}'s central
systems idea as a typed unified-source model for multi-artifact eBPF
applications, where programs, maps, loaders, pinning choices, and optional
kernel-extension code are represented before artifact boundaries are split.
Second, it analyzes how the implementation uses that model to check selected
cross-boundary invariants, including program-handle use, program signatures,
map access types, ring-buffer API use, config mutability, helper-scope
restrictions, kernel-context allocation checks, perf-event group bounds, and
stack-limit violations before load or attach time.  Third, it
implements a reproducible artifact evaluation of test coverage, example
buildability, generated-code expansion, example marker coverage, external
source-corpus feature context, a small manual external port/build/runtime
portfolio,
static rejection cases, strict verifier loading, isolated XDP attachment,
matched XDP and TC traffic runs, perf-event loader and counter workloads, a
ring-buffer event-emission workload, direct and callback-instrumented
struct\_ops checks, and a compiler-source patch that isolates one map-update
lowering mechanism.  The
scientific contribution is not a new BPF verifier or a broad
performance-equivalence result.  Instead, it is an executable case
study of which selected cross-boundary eBPF invariants this prototype exposes
and which compatibility limits still leak through the Linux BPF toolchain.  The
artifact is designed so that a
reviewer can rerun the evaluation without changing the \tool{} source tree:
all scripts, logs, derived tables, and paper macros are stored outside the
cloned repository.

\section{Background and Motivation}

An eBPF application normally contains at least two components.  The first is a
restricted kernel program, compiled for the BPF target and accepted only if the
kernel verifier can prove memory safety, bounded execution, and legal helper
usage.  The second is a userspace coordinator that loads objects, finds maps,
sets up BPF links, handles signals, and tears resources down.  The two
components communicate through maps, ring buffers, perf events, or global data.

This split creates three kinds of complexity.  \textbf{Lifecycle complexity}
arises because operations must occur in a valid order: an object must be
opened, loaded, attached, read, and detached with type-appropriate APIs.
\textbf{Representation complexity} appears when the same concept, such as a
map or a typed context field, must be represented in source C, generated
skeleton headers, userspace lookup code, and BTF-derived type declarations.
\textbf{Compatibility complexity} appears because eBPF features evolve quickly:
struct\_ops, kfuncs, dynptrs, TCX (a newer traffic-control attach path), and
scheduler extensions depend on kernel, bpftool, and libbpf versions.

The verifier is essential, but it is intentionally late in the pipeline.  A
developer may write code, compile it, generate skeletons, link a loader, and
only then receive a verifier rejection or an attach-time failure.  A DSL can
help if it moves some of these checks earlier without hiding the underlying
eBPF execution model.  For example, a language can know that a function marked
\texttt{@xdp} expects an \texttt{xdp\_md} context and returns an XDP action,
model maps as typed global variables, and distinguish userspace functions from
eBPF functions and kfunc definitions.

\section{Design Principles}

The design principles map directly to the three complexities above.  Lifecycle
complexity motivates explicit domains and lifecycle typing, representation
complexity motivates typed maps, program handles, and generated project
structure, and compatibility complexity motivates preserving the kernel
toolchain while exposing escape hatches for fast-moving kernel interfaces.  The
next two sections then show how the language realizes these principles.

\textbf{Preserve the kernel toolchain.}  \tool{} does not replace the kernel
verifier, clang, bpftool, or libbpf.  Instead it generates C and Makefiles that
flow through the same toolchain used by C/libbpf projects.  This choice gives
the prototype immediate access to BTF, skeleton generation, and existing
kernel diagnostics.  It also means compatibility failures remain visible and
measurable.

\textbf{Make domains explicit.}  The attribute system separates userspace,
eBPF, helper, and kernel-module code.  This matters because the same surface
syntax cannot mean the same thing in every domain.  Userspace can allocate and
print freely, eBPF code must obey stack and helper restrictions, and kfunc code
targets a module build.  A unified source model is useful only if the compiler
keeps these domains separate internally.

\textbf{Move verifier-relevant checks earlier.}  The compiler performs safety
analysis before code generation.  In our run, \texttt{safety\_demo.ks} is
rejected because the compiler estimates 608 bytes of stack use, above the
512-byte eBPF limit.  The check does not make the verifier redundant, but it
can reduce late failures for covered cases that would otherwise require
inspecting verifier logs.

\textbf{Expose low-level escape hatches.}  eBPF users often need new helpers,
kernel functions, or BTF-specific types before high-level libraries expose
them.  \tool{} supports includes and extern declarations so that a program can
refer to generated header declarations.  For a research prototype, this escape
hatch matters because the Linux BPF API surface changes faster than a DSL can
stabilize.

\section{\tool{} Overview}

\tool{} uses attributes to assign functions to execution domains.  A function
marked \texttt{@xdp}, \texttt{@tc}, \texttt{@probe}, \texttt{@tracepoint}, or
\texttt{@perf\_event} becomes an eBPF entry point.  A function marked
\texttt{@helper} is callable from eBPF code.  A normal \texttt{main} function
is compiled into userspace C.  A function marked \texttt{@kfunc} can trigger
kernel-module generation.  The source therefore contains the application
structure that would otherwise be scattered across multiple files.

\begin{lstlisting}[language=C,caption={Minimal shape of a unified-source eBPF application.}]
include "xdp.kh"

@xdp fn pass_all(ctx: *xdp_md) -> xdp_action {
  return XDP_PASS
}

fn main() -> i32 {
  var prog = load(pass_all)
  attach(prog, "lo", 0)
  detach(prog)
  return 0
}
\end{lstlisting}

This small example already exhibits the three relationships that the next
section makes precise: functions are assigned to execution domains by attribute,
maps are declared as typed globals shared across domains rather than as separate
C sections and userspace lookup code, and the program lifecycle
(\texttt{load}/\texttt{attach}/\texttt{detach}) is expressed over typed program
handles.  The same global-declaration pattern applies to ring buffers and perf
events: high-level operations stay explicit in source while the generated code
performs the required libbpf calls.  In the current implementation, this also
extends to selected dynptr-heavy paths: a source-level
\texttt{events.reserve()} can lower to
\texttt{bpf\_ringbuf\_reserve\_dynptr}, \texttt{bpf\_dynptr\_data}, field writes
through \texttt{bpf\_dynptr\_write}, and corresponding submit/discard helper APIs without
requiring the author to hand-write the dynptr lifetime sequence.  Helper-visible
map-value field reads can likewise lower through
\texttt{bpf\_dynptr\_from\_mem} plus \texttt{bpf\_dynptr\_data}.  All of this structure lives in one source
file that the compiler later splits into kernel C, userspace C, skeleton
references, and build logic.  Section~\ref{sec:typesystem} gives the typing
rules behind each relationship.  The same splitting mechanism also supports a
less conventional path exercised by the repository's
\texttt{examples/test\_exec.ks}: after attaching an XDP program, the generated
userspace stage can replace itself with a Python process, preserve the relevant
map file descriptors across \texttt{exec()}, and import an auto-generated
wrapper that exposes those maps as Python objects.  This is still the same
unified-source claim, just carried across a
kernel$\leftrightarrow$userspace$\leftrightarrow$script boundary instead of
stopping at a C loader.

\section{Language and Type System}
\label{sec:typesystem}

Because the paper's central claim is a \emph{typed} unified-source model, this
section specifies the parts of the type system that encode cross-boundary
structure.  We describe the model informally but precisely; we do not give a
full formal semantics or a soundness proof, and we are explicit below about what
each rule does and does not guarantee.  Every rule in this section corresponds
to a concrete diagnostic exercised by the rejection corpus in
Section~\ref{sec:eval} (the file names in parentheses are the cases in
\texttt{experiments/static\_checks/}).

\textbf{Domains as attributes.}  A \tool{} program is a flat sequence of
declarations.  A function attribute assigns each function to an execution domain
$d \in \{\textit{user}, \textit{ebpf}, \textit{helper}, \textit{kfunc}\}$:
\texttt{@xdp}, \texttt{@tc}, \texttt{@probe}, \texttt{@tracepoint}, and
\texttt{@perf\_event} mark eBPF entry points; \texttt{@helper} marks
kernel-shared eBPF code; \texttt{@kfunc} and \texttt{@private} mark
kernel-module code; an unattributed \texttt{fn} is userspace.  The type checker
tags each function with its domain and rejects calls that cross the domain
lattice illegally, for example a userspace function calling an \texttt{@helper}
(\texttt{helper\_from\_userspace}) or kernel-context allocation requested from a
userspace or XDP domain (\texttt{gfp\_from\_userspace}, \texttt{gfp\_from\_xdp}).
Each eBPF attribute also fixes a context type and a return type: \texttt{@xdp}
requires \texttt{ctx:\,*xdp\_md} and returns \texttt{xdp\_action}; \texttt{@tc}
requires \texttt{*\_\_sk\_buff} and returns \texttt{i32}.  A function whose
signature disagrees with its attribute is rejected before code generation
(\texttt{xdp\_wrong\_context}, \texttt{tc\_wrong\_context},
\texttt{xdp\_wrong\_return}, \texttt{probe\_wrong\_return}).

\textbf{Maps as typed globals.}  A map is a typed global declaration
\texttt{var m : hash<K,V>(N)} or \texttt{array<K,V>(N)}, rather than a separate C
section plus userspace lookup code.  The element types form a contract checked
at every access site: an index \texttt{m[k]} requires \texttt{k\,:\,K} and yields
a value of type \texttt{V}, in both eBPF and userspace domains.  Map accesses
with the wrong key or value type, or to an undeclared map, are rejected
(\texttt{map\_string\_key}, \texttt{map\_string\_value}, \texttt{map\_undefined}).

\textbf{Program lifecycle as typestate.}  The core cross-boundary invariant is
program lifecycle.  \tool{} distinguishes a \texttt{FunctionRef} (a reference to
an \texttt{@}-attributed function) from a \texttt{ProgramHandle} (a loaded
program), and types the lifecycle builtins so that only \texttt{load} can
produce a \texttt{ProgramHandle}:
\[
\frac{\Gamma \vdash f : \texttt{FunctionRef}}
     {\Gamma \vdash \texttt{load}(f) : \texttt{ProgramHandle}}
\qquad
\frac{\Gamma \vdash h : \texttt{ProgramHandle}}
     {\Gamma \vdash \texttt{detach}(h) : \texttt{unit}}
\]
\[
\frac{\Gamma \vdash h : \texttt{ProgramHandle} \quad
      \Gamma \vdash t : \texttt{string},\; \mathit{fl} : \texttt{u32}}
     {\Gamma \vdash \texttt{attach}(h, t, \mathit{fl}) : \texttt{u32}}
\]
Because \texttt{attach} and \texttt{detach} consume a \texttt{ProgramHandle} and
only \texttt{load} produces one, ``attach before load'' is a \emph{type} error
rather than a runtime failure: passing a bare \texttt{FunctionRef}
(\texttt{attach\_without\_load}), an integer (\texttt{attach\_integer\_handle}),
or a string (\texttt{load\_string}, \texttt{detach\_string}) to a lifecycle
builtin does not type-check.  This is a deliberately lightweight typestate
discipline.  What it guarantees is that lifecycle builtins are only applied to
values of the correct lifecycle type, so the producer/consumer relationship
between \texttt{FunctionRef} and \texttt{ProgramHandle} encodes the intended
\texttt{load}\,$\rightarrow$\,\texttt{attach}\,$\rightarrow$\,\texttt{detach}
ordering.  What it does \emph{not} guarantee is full path-sensitive typestate
soundness: the current checker does not statically rule out use-after-detach,
double-attach, or leaked handles on every control-flow path.  Establishing those
properties would require a flow-sensitive analysis we have not implemented, and
we therefore treat lifecycle typing as a research direction rather than a closed
result.

\textbf{From invariants to diagnostics.}  The value of making these relationships
typed is that each cross-boundary invariant becomes a localized compiler
diagnostic instead of an implicit naming convention or a late verifier or
attach-time failure.  Section~\ref{sec:eval} measures the current diagnostic
behavior; Section~\ref{sec:threats} is explicit that the existing corpus tests
whether these checks are wired and stable, not yet whether they catch the
mistakes developers make in practice.

\section{Implementation}

The evaluated implementation is an OCaml compiler built with dune.  It parses
\tool{} source, resolves imports and includes, builds symbol tables, performs
multi-program analysis and type checking, runs safety analysis, generates an
intermediate representation, and emits C artifacts.  For an input
\texttt{foo.ks}, the generated project normally contains \texttt{foo.ebpf.c},
\texttt{foo.c}, a \texttt{Makefile}, and, when kfuncs are present,
\texttt{foo.mod.c}.  The Makefile uses bpftool to dump \texttt{vmlinux.h},
clang to produce a BPF object, bpftool to generate a skeleton header, and gcc
to link the userspace loader with libbpf, libelf, and zlib.

The compiler's architecture aligns with the research hypothesis.  It has
separate modules for maps, userspace code generation, eBPF C generation,
kernel-module generation, context-specific code generation, safety checking,
BTF parsing, dynptr bridging, tail-call analysis, and struct\_ops registry
logic.  These modules encode eBPF concepts directly rather than treating the
language as a thin preprocessor over C.  The design therefore provides places
where domain-specific checks can be added: map type compatibility, context
field typing, stack estimates, function signature validation, and userspace
API generation.  The dynptr path is illustrative here because it bridges a
usability gap rather than only a code-size one: the eBPF backend tracks
dynptr-backed ring-buffer reservations so that source-level record writes can
be emitted as \texttt{bpf\_dynptr\_write} calls and the corresponding
\texttt{submit}/\texttt{discard} operations can target the same dynptr object.
For helper-visible map-value field reads, the backend can emit
\texttt{bpf\_dynptr\_from\_mem} plus \texttt{bpf\_dynptr\_data} instead of
requiring the developer to hand-assemble that helper sequence.
Those generated userspace APIs are not limited to C entry
points: the implementation also contains explicit support for \texttt{exec()}
handoff, including generated Python wrappers that reconstruct typed access to
inherited eBPF maps from compile-time map metadata.

\section{Evaluation Methodology}
\label{sec:eval}

We evaluate \tool{} as a compiler and systems artifact.  The goal is to test
whether the unified-source model centralizes recurring eBPF project structure,
preserves a compile/build path to ordinary libbpf artifacts on our host, and
rejects selected pre-load errors.  We do not claim
packet-rate performance equivalence in this paper, because the current traffic
evidence is limited to matched XDP and TC pass/count programs on local veth
pairs, and generated-loader runtime evidence is a perf-event lifecycle latency
check rather than a dispatch-loop throughput benchmark.  Instead, we
focus on reproducible evidence that
repository examples exercise multiple eBPF domains, selected errors are
rejected early, generated artifacts are classified by build and verifier-load
outcomes against the local toolchain, matched local workloads cover several
program classes, struct\_ops checks separate object compatibility, loopback
TCP workload behavior, and generated-skeleton compatibility, a skeleton repair
checks one version-aware generated-build fix, and one observed runtime gap can
be traced to a concrete lowering rule.

The evaluation runs on Linux \KSKernelVersion{} using clang 18, bpftool 7.7,
libbpf 1.3.0, OCaml 4.14, and dune 3.14.  The host CPU is a
\KSCPUModel{} with \KSCPUCount{} CPUs, CPU0 governor \KSCPUGovernor{}, turbo
\KSTurboState{}, and virtualization reported as \KSVirtualization{}.  The
microbenchmarks and object workloads do not pin CPUs or change the governor, so
their nanosecond and event-rate differences are not statistically interpreted
and should be read as local comparisons rather than general performance claims.
The evaluated commit is
\KSCommitShort{}.  The evaluation comprises 23 measurement scripts under
\texttt{experiments/}, organized around four paper claims.  The
generated-structure claim uses the example build matrix, matched
source-footprint proxy, external source-only scan, and a small manual external
port/build/runtime portfolio.  The early-checking claim uses unit tests and a
targeted static corpus for selected pre-load rejections.  The compatibility
claim uses strict verifier loading, isolated XDP attach/detach,
generated-loader lifecycle checks, object-level
perf-event/ring-buffer/struct\_ops workloads, and scheduler-extension
diagnostics to test local compatibility boundaries.  The runtime-behavior claim
uses BPF\_PROG\_TEST\_RUN microbenchmarks, matched XDP/TC traffic, longer traffic
stress, and lowering ablations to isolate local runtime behavior and one
map-update mechanism.  The paper reports the
compiler-source patch ablation as the canonical mechanism-isolation result
because it changes the compiler path
rather than editing generated output.  The traffic scripts compare generated
objects with hand-written C/eBPF baselines built with the same clang/libbpf
toolchain and matched for program type, map schema, attachment point, and
observable counter or byte-transfer semantics.  The baseline policy permits only
the idiomatic C/eBPF needed for those matched semantics.  The
baseline sources are tracked under \texttt{experiments/baselines/}.  Their
construction policy is to match the generated program's observable semantics
and avoid C-only shortcuts outside the map operation, attach point, event, or
socket behavior under test.  The oracles check semantic equivalence at the
boundary each workload exposes: exact map counts, perf-counter equality,
submitted/received ring-buffer events, byte-transfer completion,
load/attach/detach success, or callback flags.  The
perf-event counter,
ring-buffer, and struct\_ops scripts use one libbpf runner for both objects and
check BPF map counts, submitted/received event counts, load/attach/detach
outcomes, TCP socket byte-transfer oracles, or callback-reachability flags.
The main mechanism-isolation
script copies the \tool{}
compiler source, applies a tracked compiler-source patch for array-map
increment lowering, rebuilds the patched compiler, and reruns the same count
benchmark.

We measure seventeen outcomes across the four claims, and report each one with
its full procedure in the correspondingly grouped subsections of
Section~\ref{sec:results}; Table~\ref{tab:claimscope} gives the artifact,
baseline, oracle, and boundary per claim.  For \emph{early checking} we report
compiler regression coverage (the repository test suite) and early rejection
(static cases whose expected diagnostics are checked by the harness).  For
\emph{build, load, and compatibility} we report example buildability,
verifier-load and isolated XDP attach/detach outcomes, direct struct\_ops object
compatibility, the struct\_ops TCP-workload and callback-reachability checks, the
struct\_ops skeleton repair, the scheduler-extension verifier diagnostic and
opt-in attach/progress harness, and generated-loader lifecycle timing.  For
\emph{generated structure} we report example marker coverage, the
generated-structure expansion factor (lines of \tool{} source versus generated C
and Makefile lines), the matched source-footprint proxy, the external
source-corpus scan, and the external port/build/runtime portfolio.  For
\emph{runtime behavior} we report the map-update mechanism-isolation ablation and
runtime-sanity workloads (matched XDP/TC iperf3 traffic with a longer stress
rerun, a perf-event page-fault counter, and a ring-buffer event-emission
workload).  Two caveats apply throughout and are not repeated at each outcome:
example feature labels are derived from lexical markers in the source and are a
coarse corpus classifier rather than a feature-completeness audit, and all timing
and event-rate numbers are local, unpinned samples on the host described above
rather than statistically controlled benchmarks.

\begin{table*}[t]
\centering
\caption{Claim scope, actors, and oracles used in the evaluation.}
\label{tab:claimscope}
\scriptsize
\setlength{\tabcolsep}{3pt}
\begin{tabular}{@{}p{0.12\textwidth}p{0.18\textwidth}p{0.14\textwidth}p{0.25\textwidth}p{0.16\textwidth}@{}}
\toprule
Claim & Artifact under test & Baseline or actor & Oracle & Boundary \\
\midrule
Project generation & Repository examples lowered to generated projects & Compiler and make harness & Build rows, feature markers, generated source lines & Repository corpus only \\
Source footprint & Matched local workload sources & Hand-written C/eBPF objects and C/libbpf runners & Nonblank noncomment source-line counts with explicit exclusions & Matched source-footprint proxy, not developer time \\
External source corpus & Pinned public eBPF example/application repositories & Source-only feature classifier & File/SLOC counts, \texttt{SEC()} sections, and feature markers & No translation, build, verifier, attach, or runtime evidence \\
External port portfolio & \KSExternalPortWorkloads{} pinned \texttt{xdp-tutorial} XDP sources & Original external C/eBPF sources & Generated-Makefile build for \tool{} ports, direct clang compile for C/eBPF, XDP attach, iperf3 traffic, and map-key increment for the counter workload & Small manual XDP portfolio only, not automated or broad portability \\
Early checks & Unit tests and targeted invalid programs & Dune and static-check harnesses & Test success and expected diagnostic categories & Selected cases, not soundness \\
Compatibility & Generated objects and generated perf loader & Local kernel/toolchain & Pinned-program load criterion, attach state, loader attach/read/detach and invocation timing & Local host only \\
Runtime sanity & Generated XDP, TC, perf-counter, and ring-buffer objects & Matched C/eBPF objects or shared libbpf runner & iperf3 JSON, positive map counts, perf-counter equality, submitted/received/drop counts & Local workloads only \\
Tcp-congestion struct\_ops & Generated and C/eBPF tcp-congestion objects & Shared libbpf runner and socket workload & Load/attach/detach, loopback TCP byte-transfer, clean and loss-injected callback flags & One local tcp-congestion object and selected callback profiles only \\
Struct\_ops skeleton repair & Generated userspace skeleton projects & Installed libbpf header & Detect absent map-link field, remove incompatible assignments, rebuild & Local generated-userspace build repair only \\
Scheduler extension & Generated \texttt{sched\_ext} object and five-callback C/eBPF control & Local kernel/toolchain & \texttt{bpftool prog loadall}, pinned programs, register/workload/unregister state & Toy FIFO workload only, not scheduler policy quality \\
\bottomrule
\end{tabular}
\end{table*}

The harness treats generated data as first-class evidence.  It writes raw
stdout and stderr logs for every compiler and make invocation, CSV rows for
each example, JSON summaries for scripts, and a small TeX file containing the
numeric macros used in this paper.  This organization is deliberately close to
artifact-evaluation practice: the paper does not rely on a spreadsheet that
was manually edited after the run, and the reader can trace each aggregate
number to a row in \texttt{results/examples\_summary.csv}.  We also keep
temporary build products under \texttt{results/build} so that failures can be
inspected after the harness exits.

Source lines of code (SLOC) are counted as nonblank, noncomment lines.
Generated \texttt{vmlinux.h} and generated skeleton headers are excluded from
the expansion factor because they are bpftool outputs rather than \tool{}
compiler outputs.  The expansion therefore counts only source and build logic
produced directly by \tool{}: userspace C, eBPF C, optional kernel-module C, and
Makefile lines.  This definition is conservative for the generated-structure
claim because it omits the large skeleton header that a low-level build still
needs.

\texttt{run\_source\_footprint.py} adds a separate matched source-footprint proxy for
the local matched workloads.  It counts nonblank noncomment lines in the
maintained \tool{} application source, the hand-written C/eBPF baseline objects,
and the C/libbpf runner or loader files when that workload requires one.
Generated C, generated Makefiles, generated \texttt{vmlinux.h}, skeleton
headers, \tool{} library headers, and Python experiment harness code are
excluded.  This metric is closer to a hand-written baseline than the generated
expansion factor, but it is still not a developer-time study.

\texttt{run\_external\_corpus.py} adds an external source-only context check. It
fetches pinned commits of \texttt{libbpf-bootstrap}, \texttt{xdp-tutorial}, and
\texttt{scx}, scans selected application/example C and header paths, and records
file counts, source lines, \texttt{SEC()} sections, and feature-family markers.
The scan excludes vendored \texttt{vmlinux.h} headers, generated files, build
outputs, Rust userspace code, and repository support libraries outside the
selected application/example paths.  The scan is useful for checking whether the
local artifact exercises externally visible eBPF feature families, but it is not a
porting, build, verifier, or runtime experiment.
The script also records a manual classifier spot-check over
\KSExternalCorpusAuditSamples{} representative files.  The audit compares
expected feature markers with the lexical detector output and matches
\KSExternalCorpusAuditMatched{} of \KSExternalCorpusAuditSamples{} samples, with
\KSExternalCorpusAuditFalsePositives{} false-positive feature labels and
\KSExternalCorpusAuditFalseNegatives{} false-negative feature labels.  The audit is a
reconstructability check for the selected marker rules, not a statistical
precision/recall estimate.

\texttt{run\_external\_port.py} adds a small external port/build/runtime
portfolio.  It fetches \texttt{xdp-tutorial} commit
\texttt{\KSExternalPortCommitShort{}}, builds manually written \tool{} ports for
\texttt{basic01-xdp-pass}, \texttt{basic02-prog-by-name}, and
\texttt{basic03-map-counter} through generated Makefiles, and compiles the
corresponding original external C/eBPF sources directly with clang.  It then
attaches each object to a fresh veth in a network namespace and runs iperf3
traffic.  The oracle is deliberately semantic and narrow: all workloads must
pass traffic, and \texttt{basic03-map-counter} must also increment map key
XDP\_PASS.  The experiment does not claim automated translation,
upstream-project build coverage, exact packet-count equivalence, or broad
external application portability.

\section{Results}
\label{sec:results}

\textbf{Result map.}  The evaluation is organized by four claims.  The example
build matrix and generated-code expansion test the unified-source structure
claim.  The unit tests, static rejection corpus, and safety rejection test the
early-checking claim.  The verifier, attach, perf-event, ring-buffer, and
struct\_ops checks test local compatibility.  The microbenchmarks, compiler
patch, traffic runs, and object workloads test runtime behavior and mechanism
isolation without claiming broad performance equivalence.  The remainder of this
section presents the results in four groups that follow these claims: early
checking, build/load/attach compatibility, generated structure and source
footprint, and runtime behavior.

\subsection{Early checking}

\textbf{Test suite.}  The compiler test suite reports \KSTotalUnitSuites{}
successful suites and \KSTotalUnitTests{} passing tests with no reported
failures.  The tests cover lexer/parser behavior, type checking, maps, map
assignment, userspace code generation, context field typing, safety checking,
ring buffers, perf event attach, detach APIs, kfunc attributes, struct\_ops,
BTF parsing, and other regression cases.  This breadth is important because
the evaluation of a DSL depends on whether the implementation actually encodes
domain concepts rather than only accepting example syntax.

\textbf{Static rejection corpus.}  The static-check harness compiles
\KSStaticTotal{} targeted programs and verifies that every observed outcome
matches the expected result.  The corpus includes one positive control and
\KSStaticExpectedFailures{} expected compiler rejections:
\KSStaticLifecycle{} lifecycle API misuses, \KSStaticProgramSignature{}
program-signature violations, \KSStaticMapType{} map type or symbol errors,
\KSStaticTypeSystem{} general type-system errors, \KSStaticSymbolValidation{}
symbol-validation errors, \KSStaticConfigBoundary{} config-boundary violation,
\KSStaticHelperScope{} helper-scope violation, \KSStaticKernelContext{}
kernel-context allocation violations, \KSStaticPerfEventGroup{} perf-event
group-bound violations, \KSStaticRingbuf{} ring-buffer API misuse, and
\KSStaticSafety{} stack-limit violation.  These cases are not a
soundness proof, and we are deliberately careful about what they do and do not
show.  The corpus is author-constructed: each invalid program was written to
exercise one specific check and confirmed to be rejected.  It is therefore a
\emph{diagnostic-contract regression test}.  It demonstrates that the checks of
Section~\ref{sec:typesystem} are implemented, attached to the right error
classes, and stable, so that a future compiler change which silently accepts one
of these invalid programs, or changes its diagnostic category, will fail the
paper-number generation path.  It does \emph{not} establish that these checks
catch the mistakes developers actually make, nor that they reject errors the
C/libbpf path would only surface late at verifier-load or attach time.  That
stronger claim requires an organic mistake corpus rather than a synthetic one;
we make this gap explicit in Section~\ref{sec:threats} and treat the organic
study as the key remaining early-checking experiment.

\subsection{Build, load, attach, and compatibility}

\textbf{Example buildability.}  Table~\ref{tab:build} summarizes the example
build results.  Of \KSTotalExamples{} examples, \KSKSCompileOK{} compile from
\tool{} source.  \KSMakeOK{} then build into generated C/eBPF artifacts.  The
single compile-time rejection is \texttt{safety\_demo.ks}, where safety
analysis detects 608 bytes of stack usage and halts before C generation.  This
is a positive result for the early-checking hypothesis: the compiler catches a
verifier-relevant problem early and provides a specific repair suggestion.

\begin{table}[t]
\centering
\caption{Repository examples at commit \KSCommitShort{}: \KSMakeOK{}/\KSTotalExamples{}
build fully, with one intended safety rejection and two local
struct\_ops/libbpf skeleton mismatches.}
\label{tab:build}
\begin{tabular}{lrr}
\toprule
Outcome & Count & Fraction \\
\midrule
Examples total & \KSTotalExamples{} & 100\% \\
\tool{} compile success & \KSKSCompileOK{} & \KSKSCompilePct{} \\
Generated project build success & \KSMakeOK{} & \KSMakePct{} \\
Safety rejection & 1 & \KSSafetyPct{} \\
Struct\_ops/libbpf mismatch & 2 & \KSStructOpsMismatchPct{} \\
\bottomrule
\end{tabular}
\end{table}

The two generated build failures are \texttt{sched\_ext\_simple.ks} and
\texttt{struct\_ops\_simple.ks}.  Both compile to eBPF objects and generate
skeletons, but the generated userspace skeleton code references a
\texttt{struct bpf\_map\_skeleton.link} field not available in the installed
libbpf 1.3.0 headers.  The fast-moving struct\_ops path therefore exposes a
compatibility issue, not a parser or type-checking failure, and it shows that a
unified-source language still inherits version skew across
bpftool, libbpf, and kernel headers.

\textbf{Struct\_ops object compatibility.}  To separate that skeleton-header
failure from object compatibility, \texttt{run\_struct\_ops\_compat.py}
compiles \texttt{struct\_ops\_simple.ks} with \texttt{make ebpf-only},
compiles a hand-written C/eBPF object with the same minimal tcp-congestion
function set, and loads both with one libbpf runner.  The runner finds the
struct\_ops map, calls \texttt{bpf\_map\_\_attach\_struct\_ops}, destroys the
returned link, and records detach success.  On
this host, \texttt{bpftool} is \KSStructOpsBpftoolVersion{} but
\texttt{libbpf-dev} is \KSStructOpsLibbpfVersion{}, and the local
\texttt{struct bpf\_map\_skeleton} header reports skeleton map-link support as
\KSStructOpsSkeletonLinkSupported{}.  Without generated skeletons, both the
generated object and the C/eBPF object load, attach, and detach in
\KSStructOpsKsDetachOK{}/\KSStructOpsTrials{} and
\KSStructOpsCDetachOK{}/\KSStructOpsTrials{} trials, respectively.  This check
does not validate scheduler-extension struct\_ops, full struct\_ops signature
modeling, or workload performance, but it narrows the
\texttt{struct\_ops\_simple.ks} generated userspace build failure to a
skeleton compatibility problem for this host.

\textbf{Struct\_ops skeleton repair.}  The local
\texttt{struct bpf\_map\_skeleton} header reports skeleton map-link support as
\KSStructOpsRepairSkeletonLinkSupported{}.  When that field is absent,
\texttt{run\_struct\_ops\_skeleton\_repair.py} removes the generated skeleton
map-link assignments that point at struct\_ops links and rebuilds the two
affected generated userspace projects.  On this host, the original generated
userspace builds succeed in
\KSStructOpsRepairBaselineBuildOK{}/\KSStructOpsRepairExamples{} cases.  After
the repair, \KSStructOpsRepairBuildOK{}/\KSStructOpsRepairExamples{} build.
The script removes \KSStructOpsRepairRemovedAssignments{} assignments across
the two examples.  The result is a local generated-userspace build repair for
the observed map-link mismatch.  It does not run the repaired binaries, integrate
the decision into the compiler, or prove portability across libbpf versions.

\textbf{Struct\_ops TCP workload.}  To move beyond load/attach/detach,
\texttt{run\_struct\_ops\_workload.py} uses one runner for the generated object
and the C/eBPF object.  After attaching each tcp-congestion struct\_ops map, the
runner creates a loopback TCP client/server pair, sets the client socket's
\texttt{TCP\_CONGESTION} option to the newly registered BPF algorithm, verifies
that the socket reports the requested algorithm, transfers
\KSStructOpsWorkloadMiB{}MiB, and detaches the link.  Across
\KSStructOpsWorkloadTrials{} trials, the generated object selects the BPF
algorithm, transfers the full byte count, and detaches in
\KSStructOpsWorkloadKsOK{}/\KSStructOpsWorkloadTrials{} trials.  The C/eBPF
object does the same in
\KSStructOpsWorkloadCOK{}/\KSStructOpsWorkloadTrials{} trials.  Median
end-to-end invocation times are \KSStructOpsWorkloadKsElapsedMs{}ms and
\KSStructOpsWorkloadCElapsedMs{}ms, respectively.  The workload is a local
socket-level oracle, not a claim about production TCP performance.

\textbf{Struct\_ops callback workload.}  To check that the TCP workload enters
BPF callbacks rather than only selecting a registered algorithm,
\texttt{run\_struct\_ops\_callback\_workload.py} instruments the generated and
C/eBPF tcp-congestion objects with a five-slot BPF array map.  Each callback
sets its own flag, and the shared runner clears the map before transfer and
reads it before detach.  The runner also prints a clean-profile
\texttt{callback\_clean\_required} field, while the Python harness computes the
profile-specific oracle from the required flag list.  Across
\KSStructOpsCallbackTrials{} trials of \KSStructOpsCallbackMiB{}MiB, the
generated object completes the byte-transfer
oracle in \KSStructOpsCallbackKsOK{}/\KSStructOpsCallbackTrials{} trials and
sets both \KSStructOpsCallbackRequired{} in
\KSStructOpsCallbackKsOracleOK{}/\KSStructOpsCallbackTrials{} trials.  The
C/eBPF object matches those results in
\KSStructOpsCallbackCOK{}/\KSStructOpsCallbackTrials{} byte-transfer trials and
\KSStructOpsCallbackCOracleOK{}/\KSStructOpsCallbackTrials{} callback-oracle
trials.  The generated object sets \texttt{cong\_avoid} and
\texttt{cwnd\_event} in \KSStructOpsCallbackKsCongAvoid{} and
\KSStructOpsCallbackKsCwndEvent{} trials, respectively, matching
\KSStructOpsCallbackCCongAvoid{} and \KSStructOpsCallbackCCwndEvent{} C/eBPF
trials.  The same script then adds \KSStructOpsCallbackLossPct{} loss to the
loopback device with \texttt{tc netem} as a callback-trigger profile and repeats
\KSStructOpsCallbackLossTrials{} \KSStructOpsCallbackLossMiB{}MiB trials per
variant.  Under loss, the generated object completes
\KSStructOpsCallbackKsLossOK{}/\KSStructOpsCallbackLossTrials{} byte-transfer
trials and sets \KSStructOpsCallbackLossRequired{} in
\KSStructOpsCallbackKsLossOracleOK{}/\KSStructOpsCallbackLossTrials{} trials.
The C/eBPF object matches with \KSStructOpsCallbackCLossOK{}/\KSStructOpsCallbackLossTrials{}
byte-transfer trials and \KSStructOpsCallbackCLossOracleOK{}/\KSStructOpsCallbackLossTrials{}
callback-oracle trials.  The \texttt{undo\_cwnd} flag is not part of the loss
oracle, and it appears in \KSStructOpsCallbackKsLossUndoCwnd{} generated loss
trials and \KSStructOpsCallbackCLossUndoCwnd{} C/eBPF loss trials.  The result is
callback-reachability evidence for two local loopback profiles, not coverage of
every tcp-congestion callback or a TCP performance result.

\textbf{Verifier-load matrix.}  Build success is still weaker than kernel
acceptance, so \texttt{run\_verifier\_matrix.py} loads generated objects with
\texttt{bpftool prog loadall} without attaching them.  The script counts a load
as successful only when \texttt{bpftool} returns success and at least one BPF
program is pinned, avoiding false positives from empty program sections.  Of
\KSVerifierTotalObjects{} objects present after evaluation,
\KSVerifierLoadOK{} load successfully under this stricter criterion.  Among the
\KSVerifierBuildObjects{} objects from fully built generated projects,
\KSVerifierBuildLoadOK{} load successfully (\KSVerifierBuildLoadOKPct{}) and
\KSVerifierBuildLoadFailed{} fail: \KSVerifierReferenceLeak{} remaining
reference-ownership failure, \KSVerifierMapCreateFailed{} map creation failure,
\KSVerifierMissingBTF{} missing kfunc symbol in kernel BTF, and
\KSVerifierNoProgramPinned{} object for which no program is pinned.  Across all
generated objects, including those from generated projects that did not fully
build, the matrix records no struct\_ops argument-type rejection in this run.
The current run has \KSVerifierRejected{} other verifier rejections.  This
result narrows the compatibility claim from
``builds'' to a measured split between verifier-clean objects and generated
objects that still need backend fixes.

\textbf{Scheduler-extension attach workload.}  Scheduler extensions are more
dangerous to test than packet filters because a successful registration changes
the host scheduler.  We therefore split this check into a load-only diagnostic
and an opt-in attach workload.  \texttt{run\_sched\_ext\_verifier.py} compiles a
five-callback C/eBPF control scheduler and the generated scheduler-extension
object, then runs \texttt{bpftool prog loadall} without scheduler attachment.
Both objects verifier-load on this host: the generated object pins
\KSSchedExtKsPinnedProgramCount{} programs and the C/eBPF control pins
\KSSchedExtCPinnedProgramCount{} programs, while sched\_ext remains
\KSSchedExtStateAfter{}.

\texttt{run\_sched\_ext\_attach.py} is intentionally excluded from the default
pipeline unless the caller passes an explicit host-scheduler opt-in.  The
script requires sched\_ext to start disabled, registers each scheduler with
\texttt{bpftool struct\_ops register}, runs a bounded
\KSSchedExtAttachWorkloadSeconds{}s CPU workload for
\KSSchedExtAttachWorkloadTrials{} trials with up to
\KSSchedExtAttachWorkloadWorkers{} workers per trial, records each worker's
iteration count, requires every worker to make progress, requires sched\_ext to
remain enabled after registration and every workload trial, unregisters the
struct\_ops map, and checks that sched\_ext returns to disabled.  In the
checked-in run,
both the \KSSchedExtAttachCProgramCount{}-callback C/eBPF control and the
\KSSchedExtAttachKsProgramCount{}-callback generated object pass this oracle:
\KSSchedExtAttachOK{} of \KSSchedExtAttachVariants{} variants register, keep
sched\_ext enabled during all trials, unregister cleanly, and leave
\KSSchedExtAttachKsRejectedAfter{} rejected sched\_ext tasks for the generated
variant.  The generated scheduler reports a median
\KSSchedExtAttachKsMedianIterations{} worker-loop iterations versus
\KSSchedExtAttachCMedianIterations{} for the C/eBPF control
(\KSSchedExtAttachIterationRatio{}), with median per-trial worker-count
coefficient of variation \KSSchedExtAttachKsFairnessCV{} versus
\KSSchedExtAttachCFairnessCV{}.  The harness provides attach, progress, and
fairness sanity evidence for a toy FIFO scheduler, not a scheduling-policy
performance result.

\textbf{Isolated XDP attach matrix.}  Verifier acceptance is still weaker than
attachability.  \texttt{run\_attach\_matrix.py} therefore takes the
verifier-clean, single-section XDP subset, creates a fresh network namespace
and veth pair per object, attaches each object to the namespace-side veth with
iproute2, confirms \texttt{prog/xdp} in \texttt{ip -d link show}, detaches it,
and removes the namespace.  The matrix attaches and detaches \KSAttachOK{} of
\KSAttachEligible{} eligible objects (\KSAttachOKPct{}), with
\KSAttachFailed{} failures.  This broadens the deployment evidence beyond one
smoke-test loader, but it is not a packet-behavior or throughput test.

\subsection{Generated structure and source footprint}

\textbf{Example marker coverage.}  The examples exercise many eBPF domains.
Table~\ref{tab:features} reports attempted examples, examples that pass the
\tool{} compiler, and examples that fully build against the local toolchain.
These counts should not be read as a complete language coverage proof.  They
show that the artifact is broader than a tracing-only DSL while making the struct\_ops
compatibility gap explicit.

\begin{table}[t]
\centering
\caption{Example marker coverage in repository examples shows broad local feature markers but XDP-heavy coverage. A = attempted, KS = \tool{} compiler success, B = generated project build success.}
\label{tab:features}
\scriptsize
\begin{tabular}{lrrr@{\hspace{0.8em}}lrrr}
\toprule
Feature & A & KS & B & Feature & A & KS & B \\
\midrule
XDP & \KSFeatXDPAttempt{} & \KSFeatXDPKS{} & \KSFeatXDPBuild{} &
TC & \KSFeatTCAttempt{} & \KSFeatTCKS{} & \KSFeatTCBuild{} \\
Probe & \KSFeatProbeAttempt{} & \KSFeatProbeKS{} & \KSFeatProbeBuild{} &
Tracepoint & \KSFeatTracepointAttempt{} & \KSFeatTracepointKS{} & \KSFeatTracepointBuild{} \\
Perf & \KSFeatPerfAttempt{} & \KSFeatPerfKS{} & \KSFeatPerfBuild{} &
Maps & \KSFeatMapsAttempt{} & \KSFeatMapsKS{} & \KSFeatMapsBuild{} \\
Ringbuf & \KSFeatRingbufAttempt{} & \KSFeatRingbufKS{} & \KSFeatRingbufBuild{} &
Dynptr & \KSFeatDynptrAttempt{} & \KSFeatDynptrKS{} & \KSFeatDynptrBuild{} \\
kfunc & \KSFeatKfuncAttempt{} & \KSFeatKfuncKS{} & \KSFeatKfuncBuild{} &
Struct\_ops & \KSFeatStructOpsAttempt{} & \KSFeatStructOpsKS{} & \KSFeatStructOpsBuild{} \\
Tail-call & \KSFeatTailCallAttempt{} & \KSFeatTailCallKS{} & \KSFeatTailCallBuild{} &
Userspace & \KSFeatUserspaceAttempt{} & \KSFeatUserspaceKS{} & \KSFeatUserspaceBuild{} \\
\bottomrule
\end{tabular}
\end{table}

The coverage table also highlights a weakness in the current benchmark suite:
XDP dominates.  That skew is common in early eBPF tools because XDP examples
are easy to compile and do not require tracepoint discovery or workload
generation.  A later performance study should rebalance the corpus so that TC,
perf events, ring buffers, kfuncs, and struct\_ops each have multiple matched
programs with comparable hand-written baselines.

\textbf{Generated structure.}  For successful examples, the median
\tool{} source size is \KSMedianKSSloc{} nonblank noncomment lines.  The
median generated project size, counting userspace C, eBPF C, optional module
C, and Makefile lines but excluding generated \texttt{vmlinux.h} and skeleton
headers, is \KSMedianGeneratedSloc{} lines.  The median expansion factor is
\KSMedianGeneratedRatio{}.  These measurements show that the compiler emits a
median \KSMedianGeneratedSloc{} generated source and build lines from
\KSMedianKSSloc{} \tool{} source lines, giving generated-structure evidence
rather than a direct developer-effort result.

The metric is intentionally conservative.  It does not say that every
generated line is exactly a hand-written line saved, and it does not compare
against a ported external application corpus.  Within these bounds it shows that
\tool{} \emph{centralizes recurring eBPF project structure for this
self-authored corpus}: the large ratios are evidence of concern centralization
on examples written by the implementation authors, not a corpus-independent
claim that \tool{} source is always smaller than hand-written code.  The external
port portfolio below is the cleanest available check on that boundary, and it
shows the ratio does not transfer unchanged to independent code (the \tool{}
ports and the external eBPF program bodies have the same line count).  Examples
such as perf events have particularly high expansion factors because attaching
and reading counters requires substantial userspace coordination code relative
to the kernel program itself.

\textbf{Matched source-footprint proxy.}
The matched source-footprint script covers \KSSourceFootprintWorkloads{} local
workloads that already have hand-written C/eBPF baselines.  Counting unique
source files, those workloads use \KSSourceFootprintUniqueKSSloc{} maintained
\tool{} application lines, while the hand-written C/eBPF objects alone use
\KSSourceFootprintUniqueCEbpfSloc{} lines
(\KSSourceFootprintUniqueObjectRatio{} as many).  Including the C/libbpf runner
or loader files used by the matched nontrivial workloads raises the maintained
C/libbpf source footprint to \KSSourceFootprintUniqueCSloc{} lines
(\KSSourceFootprintUniqueRatio{} as many as the \tool{} source).  Across
per-workload rows, \tool{} is smaller in \KSSourceFootprintKSSmaller{} of
\KSSourceFootprintWorkloads{} rows, while the three rows where it is not smaller
are tiny pass or scheduler-control cases with little userspace coordination.
The result is a matched source-footprint proxy for the local workload corpus,
not a claim about programmer time or external applications.

\textbf{External source-corpus context.}
The external source-only scan covers \KSExternalCorpusRepos{} public repositories
at pinned commits, \KSExternalCorpusFiles{} selected C/header files, and
\KSExternalCorpusSLOC{} nonblank noncomment source lines.  The selected corpus
contains \KSExternalCorpusKernelFiles{} kernel-side files,
\KSExternalCorpusUserspaceFiles{} userspace files, and
\KSExternalCorpusHeaderFiles{} local headers.  The classifier observes
\KSExternalCorpusFeatureCount{} of the tracked feature families: maps appear in
\KSExternalCorpusMapsFiles{} files, XDP in \KSExternalCorpusXDPFiles{}, TC in
\KSExternalCorpusTCFiles{}, tracepoint in \KSExternalCorpusTracepointFiles{},
kprobe in \KSExternalCorpusKprobeFiles{}, perf-event in
\KSExternalCorpusPerfEventFiles{}, ring-buffer in
\KSExternalCorpusRingbufFiles{}, struct\_ops in
\KSExternalCorpusStructOpsFiles{}, scheduler-extension markers in
\KSExternalCorpusSchedExtFiles{}, and kfunc-related markers in
\KSExternalCorpusKfuncFiles{}.  These rows provide external feature-context
evidence for the paper's chosen local program classes.  They also narrow the claim: the
scan found no tail-call marker in this selected corpus, and no external program
is translated, built, verifier-loaded, attached, or run by this experiment.
A manual spot-check of the lexical classifier matches
\KSExternalCorpusAuditMatched{} of \KSExternalCorpusAuditSamples{} sampled files,
with \KSExternalCorpusAuditFalsePositives{} false-positive feature labels and
\KSExternalCorpusAuditFalseNegatives{} false-negative feature labels.

\textbf{External port/build/runtime check.}
The external port harness uses three pinned \texttt{xdp-tutorial} XDP programs
as concrete checks beyond source-only scanning: \texttt{basic01-xdp-pass},
\texttt{basic02-prog-by-name}, and \texttt{basic03-map-counter}.  The \tool{}
ports total \KSExternalPortKSSloc{} nonblank noncomment lines, while the
original external C/eBPF program bodies total \KSExternalPortCSloc{} lines.  This
parity is the most important number in the external check, and it directly
qualifies the generated-structure result: on code the \tool{} authors did not
write, the large source-footprint ratios of the self-authored corpus do not
reproduce.  The comparison is not perfectly matched---the \tool{} count includes
the userspace \texttt{main} that loads and attaches the program, whereas the
external C count is the eBPF program body alone and relies on
\texttt{xdp-tutorial}'s shared userspace loader that this count excludes---so the
honest reading is that the external port neither confirms a corpus-independent
line-count advantage nor shows \tool{} to be larger; it shows that the
centralization benefit is corpus-dependent and must be measured per workload
rather than assumed.  The
\tool{} ports build through generated Makefiles, and the original external
C/eBPF sources compile directly to BPF objects with clang.  All
\KSExternalPortVariants{} objects attach to isolated veth devices and pass
iperf3 traffic in \KSExternalPortTrials{} one-second trials.  The map-counter
workload also increments the XDP\_PASS counter key in both variants, with median
map update rates of \KSExternalPortMapKSMedianMpps{} and
\KSExternalPortMapCMedianMpps{} million updates/s for \tool{} and C/eBPF,
respectively.  Receiver-throughput medians range from
\KSExternalPortKSMedianGbpsRange{} Gb/s for the \tool{} ports and
\KSExternalPortCMedianGbpsRange{} Gb/s for the original C/eBPF objects.  These
throughput and counter-rate numbers are descriptive local samples, not a
ranking.  The check is useful external port sanity evidence, but it remains a
small manual XDP portfolio.

\subsection{Runtime behavior and mechanism isolation}

\textbf{Runtime microbenchmarks and lowering ablation.}
Table~\ref{tab:micro} uses BPF\_PROG\_TEST\_RUN for 100000 repetitions over
seven trials and reports median average nanoseconds with min--max ranges.  The
pass rows come from \texttt{run\_microbench.py}, and the count rows come from
\texttt{run\_compiler\_patch\_ablation.py}.  We intentionally use this single
ablation run for all count rows so the current, patched, and C objects share
one timing run.  The count harness resets the pinned \texttt{counts} map before
each trial and requires key 0 to equal \KSCPExpectedCount{} afterward.  The
minimal pass program has the same median in this harness: 2 instructions and
\KSMicroKsPassNs{} for \tool{} versus \KSMicroCPassNs{} for C.  The map counter
isolates a lowering choice.  The current compiler emits lookup plus update,
which produces \KSCPCurrentInstr{} instructions and \KSCPCurrentNs{}.  Applying
a compiler-source patch that lowers constant array-map increments to a checked
lookup plus \texttt{\_\_sync\_fetch\_and\_add} reduces the object by
\KSCPInstrReduction{} instructions (\KSCPInstrReductionPct{}) and
\KSCPNsReduction{}ns median (\KSCPNsReductionPct{}), matching the hand-written
C/eBPF baseline in this harness.

\begin{table}[t]
\centering
\caption{XDP microbenchmarks and compiler-patch lowering ablation. The experimental compiler patch removes the count gap in this harness.}
\label{tab:micro}
\scriptsize
\begin{tabular}{@{}llrr@{}}
\toprule
Bench. & Impl. & Instr. & ns (range) \\
\midrule
pass & \tool{} & \KSMicroKsPassInstr{} & \KSMicroKsPassNs{} \\
pass & C/eBPF & \KSMicroCPassInstr{} & \KSMicroCPassNs{} \\
count & \tool{} & \KSCPCurrentInstr{} & \KSCPCurrentNs{} \\
count & \tool{} compiler patch & \KSCPCompilerPatchInstr{} & \KSCPCompilerPatchNs{} \\
count & C/eBPF & \KSCPCInstr{} & \KSCPCNs{} \\
\bottomrule
\end{tabular}
\end{table}

\textbf{Traffic-driven XDP baseline.}  To check that the microbenchmark result
does not hide an attach-path collapse, \texttt{run\_xdp\_traffic.py} runs the
same pass and count programs over real TCP traffic on a veth pair.  Each trial
creates a fresh network namespace, attaches the XDP object to the receiver-side
veth, starts an iperf3 server in the namespace, and runs a host-side iperf3
client for \KSTrafficSeconds{}s.  The count variants additionally read the
\texttt{counts} map after each trial and require a positive count value.  Over
\KSTrafficTrials{} trials, the pass medians are \KSTrafficKsPassGbps{} Gb/s for
\tool{} and \KSTrafficCPassGbps{} Gb/s for hand-written C/eBPF.  The count
medians are \KSTrafficKsCountGbps{} Gb/s for \tool{} and
\KSTrafficCCountGbps{} Gb/s for C/eBPF.  In this local run, the count medians
differ by \KSTrafficCountOverheadPct{} with overlapping ranges.  The
corresponding XDP count-map invocation rates are
\KSTrafficKsCountMpps{} and \KSTrafficCCountMpps{} Mpps.  These numbers are
local veth/TCP measurements,
not NIC-rate claims, and the pass ranges include local run-to-run noise.  The
point of the run is narrower: it adds a traffic-driven check to the object-level
BPF\_PROG\_TEST\_RUN results.

\textbf{Traffic-driven TC baseline.}  To add a non-XDP workload, the same
traffic harness runs TC ingress pass and count programs attached with
\texttt{tc qdisc clsact} and a direct-action BPF filter.  Over
\KSTCTrafficTrials{} trials, the TC pass medians are
\KSTCTrafficKsPassGbps{} Gb/s for \tool{} and \KSTCTrafficCPassGbps{} Gb/s for
C/eBPF.  The TC count medians are \KSTCTrafficKsCountGbps{} Gb/s for \tool{}
and \KSTCTrafficCCountGbps{} Gb/s for C/eBPF.  In this local run, the count
medians differ by \KSTCTrafficCountOverheadPct{} with overlapping ranges.  The
count-map execution oracle reports TC BPF-invocation rates of
\KSTCTrafficKsCountMpps{} and
\KSTCTrafficCCountMpps{} Mpps,
respectively, and all TC trials have zero retransmits in the checked-in run.
The result is still local-host evidence, but it directly addresses the XDP-only
limitation of the earlier runtime checks.

\textbf{Longer traffic stress.}  To check whether the one-second traffic
trials hide immediate failures, \texttt{run\_traffic\_stress.py} reruns the
same XDP and TC pass/count objects for \KSTrafficStressTrials{} trials of
\KSTrafficStressSeconds{}s each.  The stress harness reports iperf3 success for
all pass/count runs and positive map-count oracles for all count variants, but
the XDP stress rows include retransmits and one low generated pass sample.  We
therefore use this run as short-run sanity evidence, not as a stability claim
or throughput ranking.  In
the count benchmarks, XDP reports \KSTrafficStressXDPKsCountGbps{} Gb/s for
\tool{} and \KSTrafficStressXDPCCountGbps{} Gb/s for C/eBPF, while TC reports
\KSTrafficStressTCKsCountGbps{} Gb/s and \KSTrafficStressTCCCountGbps{} Gb/s.
In this local run, the XDP and TC count medians differ by
\KSTrafficStressXDPCountOverheadPct{} and
\KSTrafficStressTCCountOverheadPct{}, respectively, with overlapping ranges.
The count-map invocation rates remain positive at \KSTrafficStressXDPKsCountMpps{} and
\KSTrafficStressXDPCCountMpps{} Mpps for XDP, and
\KSTrafficStressTCKsCountMpps{} and \KSTrafficStressTCCCountMpps{} Mpps for
TC.  The rerun remains a short local stress check with local-host noise.

\textbf{Perf-event loader and counter workloads.}  To check generated userspace
coordination beyond XDP attachment in one generated-loader path, the
perf-event loader script compiles
\texttt{perf\_page\_fault.ks}, compiles a matched hand-written C/libbpf loader,
and runs each binary with \texttt{sudo -n}.  Over \KSPerfLoaderTrials{} trials,
both loaders attach two perf-event programs, read page-fault and branch-miss
counters, and detach cleanly.  The generated loader's median end-to-end
invocation latency is \KSPerfLoaderKsElapsedMs{}ms, with p90
\KSPerfLoaderKsPNinetyMs{}ms, while the C/libbpf loader reports
\KSPerfLoaderCElapsedMs{}ms and p90 \KSPerfLoaderCPNinetyMs{}ms.  The corresponding
median lifecycle invocation rates are \KSPerfLoaderKsIps{} and
\KSPerfLoaderCIps{} invocations/s.  Because this timing includes process
startup, sudo, load, attach, counter read, detach, and teardown, we treat it as
generated-loader lifecycle sanity evidence rather than a dispatch-loop
throughput claim.

To add a sustained non-XDP object workload,
\texttt{run\_perf\_event\_counter.py} compiles matched page-fault counter
objects and runs both with one libbpf runner.  Each trial faults
\KSPerfCounterPages{} pages for \KSPerfCounterRounds{} rounds, then requires the
BPF map count to equal the perf counter read.  Over \KSPerfCounterTrials{}
trials, both variants report a median \KSPerfCounterKsEvents{} counted events.
The generated object reaches \KSPerfCounterKsMps{} million events/s, while the
C/eBPF object reaches \KSPerfCounterCMps{} million events/s.  In this local
run, the medians differ by \KSPerfCounterOverheadPct{} with overlapping ranges.
These results are lifecycle and page-fault counter evidence, not a broad
perf-event throughput study.

\textbf{Ring-buffer event workload.}  To exercise an event-delivery path that
uses ref-counted ring-buffer records, \texttt{run\_ringbuf\_workload.py}
compiles matched XDP objects that reserve a ring-buffer record, write a marker,
submit the record, and increment submitted/drop counters.  One libbpf runner
executes each object through BPF\_PROG\_TEST\_RUN, consumes the ring buffer, and
also records the presence of an expected generated
\texttt{bpf\_ringbuf\_reserve\_dynptr},
\texttt{bpf\_dynptr\_data}, \texttt{bpf\_dynptr\_write}, and
\texttt{bpf\_ringbuf\_submit\_dynptr} helper pattern for the source-level
\texttt{reserve}/field-write/\texttt{submit} operations.  This additional check
does not prove that every map or ring-buffer path is dynptr-backed, but it does
make the evaluated developer-facing lowering explicit instead of inferring it
from a module name.  The runner then
requires submitted events to equal received events with zero drops, bad
markers, bad return values, or runner errors.  Over \KSRingbufTrials{} trials
of \KSRingbufEvents{} events, both variants report
\KSRingbufKsSubmitted{} submitted and \KSRingbufKsReceived{} received events
with \KSRingbufKsDropped{} drops.  The generated object reaches
\KSRingbufKsMps{} million events/s, while the C/eBPF object reaches
\KSRingbufCMps{} million events/s.  In this local run, the medians differ by
\KSRingbufOverheadPct{} with overlapping ranges.  The result is object-level
ring-buffer evidence, and it does not measure generated userspace
dispatch-loop throughput.

\textbf{Generated code size.}  Among the \KSMakeOK{} successful repository
examples with disassembly, the median generated eBPF instruction count is
\KSMedianInstructions{}, the midpoint of the two middle objects.  Small
examples compile to tiny programs, while map-heavy examples generate much
larger objects.  The metric helps
choose where runtime work is needed: examples with many generated map
operations are the most likely to diverge from hand-written C.

\textbf{Execution smoke test.}  The separate smoke test compiles
\texttt{smoke\_lo.ks}, builds the generated project, and runs the generated
loader with sudo.  The loader successfully attaches an XDP pass program to
the loopback interface and detaches it:
\texttt{XDP attached to interface: lo} followed by
\texttt{XDP detached from interface index: 1}.  This test is not a benchmark,
but it validates that this minimal generated XDP loader can complete a real
load/attach/detach lifecycle on \texttt{lo} on the test host, complementing the
object-level attach matrix.

\section{Discussion}

The evaluation supports \tool{}'s central research direction, but it also
clarifies what remains unproven.  The strongest supported claim is a
generated-structure claim: one source model can generate the multi-file project
structure needed by several eBPF example classes.  The test suite, expanded
static corpus, and safety demo support a narrower checking claim: the current
compiler rejects selected lifecycle, program-signature, map-type,
ring-buffer-API, config-boundary, helper-scope, kernel-context allocation,
perf-event group-bound, symbol-validation, type-system, and stack-limit errors
before kernel loading.  The runtime evidence is deliberately smaller.  It
shows that a minimal pass program has the same instruction count and a
comparable BPF\_PROG\_TEST\_RUN median in this local harness, and that the
map-update gap is caused by a specific lowering choice, not by the unified
source model alone.  The
compiler-source patch result strengthens that mechanism claim.  The XDP attach
matrix plus XDP, TC, perf-event, ring-buffer, direct struct\_ops, and
callback-instrumented struct\_ops checks strengthen the local deployment,
generated-loader lifecycle, object-workload, object-compatibility, and
callback-reachability evidence across five program classes, and the longer
traffic rerun adds a small stress check for the XDP and TC cases.  The
scheduler-extension attach harness adds a bounded host-scheduler progress and
fairness proxy for one toy FIFO policy.  The
artifact still does not establish broad runtime equivalence with hand-written
C/libbpf.

One especially suggestive design point is the repository's \texttt{exec()}
bridge to Python.  In that path, \tool{} does not merely generate a C loader
for a BPF object; it also carries typed map declarations across a process
replacement boundary by preserving file descriptors and emitting a Python
wrapper from the same source metadata.  This matters to the paper's
cross-boundary-structure claim because the compiler, rather than the developer,
performs the bookkeeping needed to reopen the shared state on the Python side.
The checked-in example stops at direct map reads and writes, so we do not claim
a full machine-learning integration result.  But the mechanism does expose a
practical path from eBPF-collected state into the Python analysis ecosystem,
where many current data-processing and ML workflows already live, without
requiring the map schema and descriptor plumbing to be rewritten by hand.

The struct\_ops failures are instructive.  A DSL can hide coordination code, but it
cannot hide all ABI and API drift.  When bpftool-generated skeletons and
installed libbpf headers disagree, the generated userspace C fails.  A direct
libbpf runner can load, attach, and detach the generated tcp-congestion
struct\_ops object on this host, and a loopback TCP socket can select the
registered BPF congestion-control algorithm and transfer data.  A
callback-flag variant shows that this workload reaches \texttt{cong\_avoid}
and \texttt{cwnd\_event} in both generated and C/eBPF objects, and the
loss-injected profile also reaches \texttt{ssthresh} and \texttt{set\_state}.
Thus, the measured failure is not simply an invalid object, unusable TCP
algorithm, or entirely unreached callback path.  The
skeleton-repair experiment shows that one local map-link mismatch can be
handled with a version-aware generated-header repair, but a production compiler
would still need to integrate that decision into generation, dependency checks,
or containerized toolchains.  The scheduler-extension experiments add a sharper
boundary: a \texttt{BPF\_PROG}-wrapped C/eBPF sched\_ext object and the
generated sched\_ext object both verifier-load, and an opt-in harness can
register both toy schedulers, run a bounded CPU workload, and unregister them
cleanly on this host.  This removes the previous load-only gap for the toy
scheduler, but it is still not a scheduler-performance or production-policy
study.
From a
systems-research perspective, this is part of the contribution: the artifact
exposes where a unified model centralizes structure and where the underlying
Linux BPF stack still leaks through.

The expansion-factor and source-footprint metrics should also be interpreted
carefully.  Generated C is not automatically better than hand-written C.  It
can be verbose, unidiomatic, or harder to debug, so the practical benefit
depends on diagnostic quality and source-level mapping from generated errors
back to \tool{} source.  The matched source-footprint proxy improves on a
generated-only expansion factor, but it still uses local baselines and line
counts rather than a controlled developer study.  Future work should therefore
measure time to fix injected errors, quality of compiler messages, and whether
developers can understand generated verifier failures.
The external corpus scan should be interpreted just as carefully: it measures
feature overlap in pinned public source trees, not whether \tool{} can compile
those programs or preserve their behavior.
The external port experiment improves that gap for a small pinned XDP portfolio
by adding compile/build/attach/runtime evidence, but it is still a manual
same-repository port set and cannot justify a broad external-application
portability claim.

\section{Threats to Validity}
\label{sec:threats}

\textbf{No developer-time study.}  The source-footprint proxy compares
\tool{} source against the hand-written C/eBPF baselines and C/libbpf runners
used by the local matched workload corpus.  This proxy is more direct than
comparing \tool{} input with generated output, but it still measures lines rather than
task time, debugging effort, review burden, or expert familiarity.  An expert
could write smaller C for some examples, especially tiny XDP pass/drop programs.
Conversely, generated code omits comments and may be denser than maintainable
production C.  The right interpretation is therefore ``matched source footprint
centralized for these local workloads,'' not ``developer time saved.''

\textbf{Repository examples may be biased.}  The corpus comes from the
\tool{} artifact repository, so it naturally reflects the features the implementation
authors wanted to demonstrate.  The corpus is useful for an artifact study because it
is reproducible and covers many features, but it is not a representative sample
of all eBPF applications.  A stronger corpus would include programs from
Cilium, bcc/libbpf-tools, production XDP filters, scheduler extensions, and
observability agents, then reimplement each in \tool{} and C/libbpf.  The
external source scan adds independent feature context, but it is source-only and
does not remove this threat.  The external port portfolio lowers the risk for
three small XDP tutorial workloads but leaves the broader corpus threat in
place.  This bias bears directly on the generated-structure and source-footprint
results: the large expansion and footprint ratios are measured on
author-written examples, and on the one independent workload set the \tool{} and
external eBPF line counts are equal.  We therefore scope the centralization claim
to ``concern centralization for this corpus'' and do not claim a
corpus-independent reduction in source size; establishing the latter requires the
broader reimplement-in-both study described above.

\textbf{Static checks are targeted and synthetic, not organic or complete.}
The static rejection corpus is an author-constructed regression test for
selected compiler checks, not a formal proof of lifecycle soundness or memory
safety, and not a sample of real developer mistakes.  Because each program was
written to be rejected, the corpus shows that the checks of
Section~\ref{sec:typesystem} are wired and stable, but it cannot by itself show
that early checking pays off in practice: it does not cover all invalid attach
orders, helper contracts, context-dependent verifier failures, or semantic
equivalence with hand-written C/libbpf, and it cannot quantify how often \tool{}
moves a real failure earlier than the C/libbpf path.  The experiment that would
settle this is an \emph{organic} early-checking study, which we have not yet run:
collect real eBPF mistakes (for example from the revision history and issue
trackers of \texttt{xdp-tutorial}, \texttt{libbpf-bootstrap}, and Cilium, or via
mutation of working programs), port each to \tool{}, and report the failure
stage in each toolchain---\tool{} compile-time rejection versus C/libbpf
clang-compile, verifier-load, or attach-time failure.  Such a study would
convert the current regression test into a measured claim about which late
failures \tool{} actually shifts earlier; until it is done, the early-checking
contribution should be read as ``these checks exist and are stable,'' not ``these
checks catch the bugs developers hit.''

\textbf{The compiler patch is narrow and experimental.}  The patch used in the
ablation is a tracked compiler-source patch applied to a copied compiler tree,
not an upstream change in the evaluated \tool{} commit.  The patch is
intentionally limited to constant increments on integer array maps, where lookup followed by
in-place atomic add preserves the count benchmark's behavior.  Broader claims
would require upstreaming or otherwise integrating the rule and rerunning
verifier, instruction-count, and runtime checks across hash, per-CPU, and
structured map values.

\textbf{Local traffic is not broad workload validation.}  The verifier matrix
loads generated eBPF objects and therefore catches failures beyond C
compilation.  The attach matrix goes further for verifier-clean single-section
XDP objects by installing and removing them on isolated veth devices, and the
traffic benchmarks send TCP through matched XDP and TC pass/count programs.
The generated perf-event loader test adds one repository-example loader that
opens perf events, attaches BPF programs, reads counters, detaches cleanly, and
records end-to-end invocation latency against a C/libbpf loader.  The
perf-event counter and ring-buffer benchmarks add sustained object workloads,
but they use a shared libbpf runner and therefore do not measure
generated-dispatch-loop throughput.
The struct\_ops compatibility checks also use shared libbpf runners.  The TCP
workload validates socket-level algorithm selection and byte transfer through
the generated and C/eBPF tcp-congestion objects, and the callback-flag workload
validates that a clean loopback transfer reaches \texttt{cong\_avoid} and
\texttt{cwnd\_event} in both variants.  Its loss-injected profile also reaches
\texttt{ssthresh}, \texttt{cong\_avoid}, \texttt{set\_state}, and
\texttt{cwnd\_event}.  The skeleton repair validates a local
generated-userspace build fix for the observed map-link mismatch.  The
scheduler-extension attach harness registers the generated and C/eBPF toy FIFO
schedulers and runs a short CPU workload under sched\_ext before unregistering
them.  These checks do not measure production TCP performance, cover every
tcp-congestion-control callback, evaluate scheduler-policy quality, prove
compatibility across libbpf versions, or measure generated-loader behavior.
However, the traffic runs, including the \KSTrafficStressSeconds{}s stress
rerun, are local-host veth evidence, not NIC-rate performance or long-duration
soak tests.  A complete
deployment study would run each example in a VM with known kernel configuration,
representative workloads, and explicit cleanup checks after detach.

\textbf{Kernel version skew matters.}  The struct\_ops failures show that
results depend on the alignment between kernel headers, bpftool, generated
skeleton code, and libbpf headers.  The repair experiment covers the local
map-link mismatch, but it is not a substitute for broader version testing.  The
threat is not incidental: eBPF programming is tightly coupled to kernel
versions.  A production \tool{} compiler should record the target kernel ABI,
check libbpf feature support before generation, and fail early with a
dependency diagnostic when a generated feature cannot be compiled locally.

\section{Related Work}

\textbf{eBPF and XDP foundations.}  XDP established the in-kernel programmable
packet-processing model that \tool{} targets, executing verified eBPF at the
earliest receive point~\cite{xdp}, and the broader eBPF/XDP design space and its
application classes have since been surveyed in
detail~\cite{ebpfsurvey,linux-ebpf}.  C/libbpf remains the reference workflow:
it exposes the full API surface and integrates BTF and Compile Once, Run
Everywhere (CO-RE), but it requires developers to manage kernel code, userspace
loaders, skeletons, maps, and build logic as separate
artifacts~\cite{libbpf}.  \tool{} keeps this toolchain underneath but moves the
multi-artifact structure into one type-checked source.

\textbf{Verification and optimization of kernel extensions.}  A separate line of
work hardens or optimizes the eBPF programs themselves.  PREVAIL replaces parts
of the in-kernel verifier with an abstract-interpretation analysis that is more
precise on larger programs~\cite{prevail}, and K2 uses program synthesis to
produce smaller, safe eBPF objects from existing ones~\cite{k2}.  hXDP compiles
eBPF for execution on FPGA NICs~\cite{hxdp}.  These systems operate on the
eBPF \emph{bytecode} after it exists; \tool{} is complementary and earlier in
the pipeline, generating the bytecode (and the surrounding loader and build) and
moving selected checks ahead of the verifier rather than replacing or
re-optimizing it.  Our compiler-source ablation
(Section~\ref{sec:eval}) touches the same instruction-lowering concerns that K2
optimizes, but as a single hand-identified rule, not an optimizer.

\textbf{Languages and bindings for eBPF.}  bpftrace offers a concise high-level
language but deliberately narrows the target to tracing, excluding XDP, TC,
struct\_ops, and kfunc workflows~\cite{bpftrace}.  Python-facing systems are also
important context here: BCC makes eBPF substantially easier to script from
Python and is especially influential for tracing, observability, and rapid
tooling, but its usual model still places Python as the userspace controller
around embedded or separate BPF code rather than making both sides one typed
source artifact~\cite{bcc}.  This Python-facing direction also appears in
application-level work on AI observability: eInfer uses eBPF to provide
transparent, fine-grained tracing for distributed LLM inference, reconstructing
per-request behavior across inference components without requiring source-level
instrumentation of the target serving stack~\cite{einfer-ebpf}.  That line of
work uses eBPF to \emph{observe} AI systems.  By contrast, \tool{}'s
\texttt{exec()}-to-Python path targets a different boundary: it brings the
glue between eBPF-resident state and a Python analysis consumer into the
compiler's unified-source model, so that map schemas, descriptor handoff, and
generated wrappers are derived from one typed source rather than assembled as
separate integration code.  For AI observability systems such as eInfer, this
kind of unified-source glue offers a path to reduce handwritten coordination
code between eBPF-resident state and Python-side analysis logic.  Host-language
libraries such as Aya (Rust) and
cilium/ebpf (Go) likewise improve userspace integration and type safety within
their host language, but preserve the split between the eBPF program and the
controlling application~\cite{aya,cilium-ebpf}.  \tool{} differs on both axes:
it is compiled and multi-domain rather than tracing-only, and it unifies the
program and its controller in one source rather than binding a separately
written program from a host language.

\textbf{Domain-specific languages for packet and dataplane processing.}  The
idea of compiling one typed description into a split low-level implementation has
strong precedent outside eBPF.  P4 compiles a single protocol-independent
program into target-specific parser and match-action
pipelines~\cite{p4}, and the Click modular router composes packet processing
from typed elements~\cite{click}.  \tool{} attacks an analogous but distinct
unification problem: unlike a P4 pipeline, an eBPF application must also generate
its userspace coordinator and survive a late, general-purpose kernel verifier
and fast-moving libbpf/bpftool ABIs, which is the source of the compatibility
limits we measure.

\textbf{Typestate and resource-lifecycle typing.}  \tool{}'s lifecycle discipline
(Section~\ref{sec:typesystem}) is a lightweight instance of typestate, the idea
that an object's legal operations depend on its current
state~\cite{typestate}.  We enforce only the producer/consumer handle
distinction rather than a full flow-sensitive typestate analysis, so this
connection is a design influence and a direction for future work rather than a
claim of equivalent guarantees.

\textbf{Positioning.}  \tool{} therefore occupies a point not covered by the
above: a compiled, multi-domain, single-source language whose research question
is whether eBPF application structure can be made explicit enough for a compiler
to generate the low-level project \emph{while preserving} the existing Linux BPF
toolchain.  The evaluated implementation is best read as both an implementation
and a concrete design artifact for that point in the space.

\section{Conclusion}

\tool{} is an early but useful prototype for studying typed unified-source
eBPF development.  The checked-in artifact supports three bounded claims.  First,
the evaluated compiler can centralize recurring project structure for this
repository corpus: \KSKSCompileOK{} of \KSTotalExamples{} examples compile,
\KSMakeOK{} build, and the median successful example expands from
\KSMedianKSSloc{} \tool{} source lines to \KSMedianGeneratedSloc{} generated
source and build lines.  A matched source-footprint proxy adds
\KSSourceFootprintWorkloads{} local workload rows in which unique maintained
\tool{} sources total \KSSourceFootprintUniqueKSSloc{} lines versus
\KSSourceFootprintUniqueCSloc{} hand-written C/libbpf lines when baseline
runners are included.  A pinned external source-corpus scan covers
\KSExternalCorpusFiles{} files across \KSExternalCorpusRepos{} public eBPF
repositories and observes \KSExternalCorpusFeatureCount{} tracked feature
families, giving source-only context for the local evaluation scope.  A small
manual external portfolio from \texttt{xdp-tutorial} also builds through
\tool{}'s generated Makefiles and runs alongside directly compiled original
external C/eBPF objects over \KSExternalPortTrials{} traffic trials.
Second,
the implementation exposes selected errors
before kernel loading: the targeted static corpus includes
\KSStaticExpectedFailures{} expected invalid programs, and the evaluation keeps
those diagnostics executable.  Third, the generated artifacts are not merely
syntactic outputs on this host: verifier loading, isolated XDP attachment,
matched XDP/TC, perf-event, ring-buffer, and tcp-congestion struct\_ops checks
exercise local load, attach, event, traffic, and callback paths against explicit
oracles, and the scheduler-extension harness exercises one bounded
host-scheduler attach/workload path.

The contribution is therefore a reproducible prototype study rather than a
finished eBPF replacement.  It shows how a compiler-checked unified programming
model can make selected cross-boundary eBPF structure explicit while still
lowering through libbpf and bpftool.  The remaining gates are also clear:
version-aware skeleton handling needs upstream integration and broader version
coverage, scheduler-extension support now has only a toy bounded workload
rather than policy or performance evidence, broader callback-level TCP behavior
remains incompletely tested, generated-loader dispatch-loop throughput needs
stronger workloads, and the experimental map-update compiler patch needs
upstream integration plus broader semantic and runtime validation.

\bibliographystyle{plain}
\bibliography{references}

\end{document}
